Estimation of Spatiotemporal Integro-Difference Equations for Dynamical Neural Field Models

### 2009-11-12 ltx/BrainIDE.tex
Fixed a couple of typos in intro

Until now, estimation of local connectivity structure has not been attempted. Recently it has been shown that it is possible to estimate local connectivity of systems governed by integro-difference equations by formulating a state-space model~\cite{Dewar2009}. In addition, the appropriate model selection tools have been developed~\cite{Scerri2009} allowing for the application of this technique to neural fields. In this paper we demonstrate how intracortical connectivity can be inferred from EEG data, based on a variant Jansen and Ritt neural field model.

@@

\section{Neural Field Model}
In this section we describe a variant of the Jansen and Ritt neural field model~\cite{Jansen1995} that will be used to create our connectivity estimator. This model has shown to be descriptive of a range of neurodynamics of the cortex such as evoked potentials and epileptic behavior and is capable of generating complex patterns of activity such as Turing patterns, spirals and traveling oscillations. The general form of the model is described here. The post-synaptic potentials generated at a neuronal population $s$ by action potentials arriving from all other connected populations at position $r$ can be described by
\begin{equation}\label{Eq1}
v\left( {s,t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)J\left( {r,t'} \right)dt'}.
\end{equation}
The post-synaptic response kernel $h(t)$ is described by
\begin{equation}\label{Eq2}
h(t) = \left\{ {\begin{array}{*{20}{c}}
   {\exp ( - {t \mathord{\left/
 {\vphantom {t \tau }} \right.
 \kern-\nulldelimiterspace} \tau })} & {t \ge 0}  \\
   0 & {t < 0}  \\
\end{array}} \right.,
\end{equation}
where $\tau$ is the synaptic time constant and typically has values from 10-20~ms (\textcolor[rgb]{1.00,0.00,0.00}{refs}). The synaptic response kernel can be more elaborate, using a rise and fall time constant, however this simple form serves our purposes. Equation~\ref{Eq1} relates average firing rate of a presynaptic neural population to the mean membrane potential of the post-synaptic neurons. The mean firing rate, $J$, is convolved with the response kernel to give membrane potential since the firing rate is assumed to follow a (homogeneous?) Poisson process. Nonlocal interactions between cortical populations are described by	
\begin{equation}\label{Eq3}
J\left( {s,t} \right) = \int_\Omega  {W\left( {s,r} \right)P\left( {v\left( {r,\bar t} \right)} \right)dr}  + E\left( {s,t} \right)
\end{equation}
where $\bar t = t - \frac{{\left| {s - r} \right|}}{d}$, $P$ is a mean pulse rate of population at $y$, $W$ is the connectivity kernel, $E$ is an external input, $\Omega$ is the spatial domain and $d$ is the propagation velocity from $r$ to $s$.

>>>>

Until now, estimation of local intracortical connectivity structure has not been attempted. Recently it has been shown that it is possible to estimate local coupling of systems governed by integro-difference equations by formulating a state-space model~\cite{Dewar2009}. In addition, the appropriate model selection tools have been developed~\cite{Scerri2009} allowing for the application of this technique to neural fields. In this paper we demonstrate for the first time how intracortical connectivity can be inferred from EEG data, based on a variant of the  Wilson and Cowan neural field model~\cite{Wilson1973}. This work provides a fundamental link between the theoretical advances in neural field modeling and patient specific data.

@@

\section{Neural Field Model}
In this section we describe a variant of the  Wilson and Cowan neural field model~\cite{Wilson1973} that will be used to create our connectivity estimator. This model has shown to be descriptive of a range of neurodynamics of the cortex such as evoked potentials and epileptic behavior and is capable of generating complex patterns of activity such as Turing patterns, spirals and traveling oscillations. The general form of the model is described here. The post-synaptic potentials generated at a neuronal population $s$ by action potentials arriving from all other connected populations at position $r$ can be described by
\begin{equation}\label{Eq1}
v\left( {s,t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)J\left( {r,t'} \right)dt'}.
\end{equation}
The post-synaptic response kernel $h(t)$ is described by
\begin{equation}\label{Eq2}
h(t) = \left\{ {\begin{array}{*{20}{c}}
   {\exp ( - {t \mathord{\left/
 {\vphantom {t \tau }} \right.
 \kern-\nulldelimiterspace} \tau })} & {t \ge 0}  \\
   0 & {t < 0}  \\
\end{array}} \right.,
\end{equation}
where $\tau$ is the synaptic time constant and typically has values from 10-20~ms (\textcolor[rgb]{1.00,0.00,0.00}{refs}). The synaptic response kernel can be more elaborate, using a rise and fall time constant, however this simple form serves our purposes. Equation~\ref{Eq1} relates average firing rate of a presynaptic neural population to the mean membrane potential of the post-synaptic neurons. The mean firing rate, $J$, is convolved with the response kernel to give membrane potential since the firing rate is assumed to follow a (homogeneous?) Poisson process. Nonlocal interactions between cortical populations are described by	
\begin{equation}\label{Eq3}
J\left( {s,t} \right) = \int_\Omega  {W\left( {s,r} \right)P\left( {v\left( {r,\bar t} \right)} \right)dr}  + E\left( {s,t} \right)
\end{equation}
where $\bar t = t - \frac{{\left| {s - r} \right|}}{d}$, $P$ is a mean pulse rate of population at $y$, $W$ is the connectivity kernel, $E$ is an external input, $\Omega$ is the spatial domain and $d$ is the propagation velocity from $r$ to $s$.
